---
title: "Paper_MEPOP_1"
author: "Pedro Fierro"
format: html
editor: visual
---

# PAPER 1

## The Impact of Political Campaigns on Affection: The Case of the First Chilean Constitutional Process (2022)

Authors: Pedro Fierro, Claudia Ramírez, Andrés Sherman y Sebastián Rivera

† Este documento es para compartir ideas y tener acceso al código que estamos ocupando.

## Environment

### Preparing our environment (loading libraries).

```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
rm(list=ls())
library(haven)
library(knitr)
library(lattice)
library(tidyverse)
library(here)
library(flextable)
library(devtools)
library(lavaan)
library(ggplot2)
library(plm)
library(naniar)
library(purrr)
library(psych)
library(interactions)
library(semPlot)
```

## Cross-Sectional Approach

### Importing data and creating variables

```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
#Import Data

data_w1 <- read_sav("Data_W1.sav")

# ID
data_w1$id <- data_w1$CodPanelista

# Age
data_w1$age_num <- data_w1$age

# Socioeconomic Status
data_w1$ses <- data_w1$RECO_NSE

# Education
data_w1$educ <- data_w1$P60

# Sex (1=women)
data_w1 <- data_w1%>%
  mutate(sex = ifelse(SEX == 2, 1,
                      ifelse(SEX == 1, 0, NA)))

# Ideology
data_w1$ideology <- ifelse(data_w1$P32 == 99, NA, data_w1$P32)

# Ideological Polarization
# Compute polarization as distance from the center (5), then add 1 to match desired scale
data_w1$ideo_polarization <- abs(as.numeric(data_w1$ideology) - 5) + 1

# Affective Polarization

# List of variable names
p44_vars <- paste0("P44_", 1:8)

# Loop through and create new affect_polarization variables
for (i in 1:8) {
  original_var <- data_w1[[p44_vars[i]]]
  new_var_name <- paste0("affect_polarization", i)
  
  data_w1[[new_var_name]] <- ifelse(
    as.numeric(original_var) == 99,
    NA,
    abs(as.numeric(original_var) - 5) + 1
  )
}



# Online Political Efficacy
data_w1$ope1 <- data_w1$P59_1
data_w1$ope2 <- data_w1$P59_2
data_w1$ope3 <- data_w1$P59_3
data_w1$ope4 <- data_w1$P59_4

# External Political Efficacy (recode)
data_w1$extef1 <- data_w1$P58_1
data_w1$extef2 <- data_w1$P58_2
data_w1$extef3 <- data_w1$P58_3

# To recode efficacies (intef1, intef3, extef1, extef3, extef4)
data_w1 <- data_w1 %>%
  mutate(across(c(extef1, extef2, extef3), ~ 6 - .x))

# Internal Political Efficacy
data_w1$intef1 <- data_w1$P58_4
data_w1$intef2 <- data_w1$P58_5
data_w1$intef3 <- data_w1$P58_6

# Media Exposure
data_w1$tv <- data_w1$P4_1
data_w1$cable <- data_w1$P4_2
data_w1$newspaper <- data_w1$P4_3
data_w1$radio <- data_w1$P4_4
data_w1$tradonline <- data_w1$P4_5
data_w1$online <- data_w1$P4_6
data_w1$podcast <- data_w1$P4_7
data_w1$officialsm <- data_w1$P4_8

# Social Media Exposure
data_w1$fb <- ifelse(data_w1$P5_1 == 99, NA, data_w1$P5_1)
data_w1$insta <- ifelse(data_w1$P5_2 == 99, NA, data_w1$P5_2)
data_w1$twitter <- ifelse(data_w1$P5_3 == 99, NA, data_w1$P5_3)
data_w1$whatsapp <- ifelse(data_w1$P5_4 == 99, NA, data_w1$P5_4)
data_w1$youtube <- ifelse(data_w1$P5_5 == 99, NA, data_w1$P5_5)
data_w1$tiktok <- ifelse(data_w1$P5_6 == 99, NA, data_w1$P5_6)
data_w1$discord <- ifelse(data_w1$P5_7 == 99, NA, data_w1$P5_7)
data_w1$twitch <- ifelse(data_w1$P5_8 == 99, NA, data_w1$P5_8)

# Franja Exposure
data_w1$franja <- data_w1$P6_1

# Social Media Political Use
data_w1$use1 <- data_w1$P25_5
data_w1$use2 <- data_w1$P25_6
data_w1$use3 <- data_w1$P25_7
data_w1$use4 <- data_w1$P25_8
data_w1$use5 <- data_w1$P25_9
data_w1$use6 <- data_w1$P25_10
data_w1$use7 <- data_w1$P25_11

# Interest
data_w1$polint <- data_w1$P21
data_w1$procint <- data_w1$P22
data_w1$plebint <- data_w1$P23
```

### Checking the new variables

```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
describe(data_w1$age_num)
table(data_w1$ses)
table(data_w1$educ)
table(data_w1$SEX)
table(data_w1$sex)
table(data_w1$ideology)
table(data_w1$ideo_polarization)

table(data_w1$affect_polarization1)
table(data_w1$affect_polarization2)
table(data_w1$affect_polarization3)
table(data_w1$affect_polarization4)
table(data_w1$affect_polarization5)
table(data_w1$affect_polarization6)
table(data_w1$affect_polarization7)
table(data_w1$affect_polarization8)

table(data_w1$ope1)
table(data_w1$ope2)
table(data_w1$ope3)
table(data_w1$ope4)

table(data_w1$extef1)
table(data_w1$extef2)
table(data_w1$extef3)

table(data_w1$intef1)
table(data_w1$intef2)
table(data_w1$intef3)

table(data_w1$tv)
table(data_w1$cable)
table(data_w1$newspaper)
table(data_w1$radio)
table(data_w1$tradonline)
table(data_w1$online)
table(data_w1$podcast)
table(data_w1$officialsm)

table(data_w1$fb)
table(data_w1$insta)
table(data_w1$twitter)
table(data_w1$whatsapp)
table(data_w1$youtube)
table(data_w1$tiktok)
table(data_w1$discord)
table(data_w1$twitch)

table(data_w1$franja)

table(data_w1$polint)
table(data_w1$procint)
table(data_w1$plebint)


```


```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}

## Creating Factors

### Reliability Test

cronbach_affect_polarization <- alpha(na.omit(data_w1[c("affect_polarization1", "affect_polarization2", "affect_polarization3", "affect_polarization4", "affect_polarization5", "affect_polarization6", "affect_polarization7", "affect_polarization8")]))
cronbach_affect_polarization

cronbach_ope <- alpha(na.omit(data_w1[c("ope1", "ope2", "ope3", "ope4")]))
cronbach_ope

cronbach_intef <- alpha(na.omit(data_w1[c("intef1", "intef2", "intef3")]))
cronbach_intef

cronbach_extef <- alpha(na.omit(data_w1[c("extef1", "extef2", "extef3")]))
cronbach_extef

cronbach_media <- alpha(na.omit(data_w1[c("tv", "cable", "newspaper", "radio", "tradonline", "online", "podcast", "officialsm")]))
cronbach_media

cronbach_social <- alpha(na.omit(data_w1[c("fb", "insta", "twitter", "whatsapp", "youtube", "tiktok", "discord", "twitch")]))
cronbach_social

cronbach_interest <- alpha(na.omit(data_w1[c("polint", "procint", "plebint")]))
cronbach_interest

cronbach_poluse <- alpha(na.omit(data_w1[c("use1", "use2", "use3", "use4", "use5", "use6", "use7")]))
cronbach_poluse
```


```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}

### Confirmatory Factor Analysis

data_w1_na <- na.omit(data_w1[c("id", "polint", "procint", "plebint", "ope1", "ope2", "ope3", "ope4", "intef1", "intef2", "intef3", "extef1", "extef2", "extef3", "tv", "cable", "newspaper", "radio", "tradonline", "online", "podcast", "officialsm","fb", "insta", "twitter", "whatsapp", "youtube", "tiktok", "franja", "use1", "use2", "use3", "use4", "use5", "use6", "use7", "age_num", "ses", "sex", "educ", "ideo_polarization", "affect_polarization1", "affect_polarization2", "affect_polarization3", "affect_polarization4", "affect_polarization5", "affect_polarization6", "affect_polarization7", "affect_polarization8")])
sum(is.na(data_w1_na))

data_w1_na <- data_w1_na %>%
  mutate(across(where(is.labelled), as.numeric))

cfa.model <- 'ope =~ ope1 + ope2 + ope3 + ope4
              intef =~ intef1 + intef2 + intef3
              extef =~ extef1 + extef2 + extef3
              media =~ tv + cable + newspaper + radio + tradonline + online + podcast + officialsm
              social =~ fb + insta + twitter + whatsapp + youtube + tiktok
              interest =~ polint + procint + plebint
              poluse =~ use1 + use2 + use3 + use4 + use5 + use6 + use7
              affect_polarization =~ affect_polarization1 + affect_polarization2 + affect_polarization3 + affect_polarization4 + affect_polarization5 + affect_polarization6 + affect_polarization7 + affect_polarization8'


fit_cfa <- cfa(cfa.model, data = data_w1_na)
latent_scores <- predict(fit_cfa)

data_w1_scores <- cbind(data_w1_na, latent_scores)
```

## Cross-sectional Analysis


### OLSE for Online Political Efficacy

Media Exposure on OPE

```{r echo=FALSE, message=FALSE, warning=FALSE}
ols_ope <- lm(ope ~ age_num + ses + educ + sex + interest + media + social + franja + poluse + ideo_polarization + affect_polarization, data = data_w1_scores)
summary(ols_ope)

ols_intef <- lm(intef ~ age_num + ses + educ + sex + interest + media + social + franja + poluse + ideo_polarization + affect_polarization, data = data_w1_scores)
summary(ols_intef)

ols_extef <- lm(extef ~ age_num + ses + educ + sex + interest + media + social + franja + poluse + ideo_polarization + affect_polarization, data = data_w1_scores)
summary(ols_extef)

```




```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
# Update the extraction function to avoid spaces in names
get_coef <- function(model, model_name) {
  coefs <- summary(model)$coefficients
  data.frame(
    Term = rownames(coefs),
    Estimate = coefs[, "Estimate"],
    Std_Error = coefs[, "Std. Error"],  # Rename to 'Std_Error'
    Model = model_name
  )
}

# Create a mapping of old variable names to new names
name_mapping <- c(
  ses = "Socioeconomic Status",
  sex = "Sex",
  age_num = "Age",
  educ = "Education",
  media = "Traditional Media for Elect. Info.",
  social = "Social Media for Elect. Info.",
  franja = "Franja Electoral",
  interest = "Pol. Interest",
  extef = "Ext. Efficacy",
  intef = "Int. Efficacy",
  ope = "Onl. Efficacy",
  poluse = "Political Use of Social Media",
  ideo_polarization = "Ideological Polarization",
  affect_polarization = "Affective Polarization"
)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Assuming coef_df is already prepared as per previous instructions

# Apply the function to both models
coef_ope <- get_coef(ols_ope, "OPE")
coef_intef <- get_coef(ols_intef, "IPE")
coef_extef <- get_coef(ols_extef, "EPE")

# Combine the data frames and remove the intercept
coef_efficacies <- rbind(coef_ope, coef_intef, coef_extef)
coef_efficacies <- coef_efficacies[!coef_efficacies$Term %in% c("(Intercept)"), ]


# Calculate 95% confidence intervals and determine significance
coef_efficacies <- coef_efficacies %>%
  mutate(
    Lower_CI = Estimate - 1.645 * Std_Error,  # Lower bound of the CI
    Upper_CI = Estimate + 1.645 * Std_Error,  # Upper bound of the CI
    Significant = if_else(Lower_CI > 0 & Upper_CI > 0 | Lower_CI < 0 & Upper_CI < 0, TRUE, FALSE)  
    )


# Apply the mapping to the dataframe
coef_efficacies <- coef_efficacies %>%
  mutate(Term = factor(Term, levels = names(name_mapping), labels = name_mapping))

# Define a dodge width for better separation
dodge_width <- 0.5

# Create the plot with updated names
ggplot(coef_efficacies, aes(x = Term, y = Estimate, ymin = Lower_CI, ymax = Upper_CI, color = Model, group = Model)) +
  geom_pointrange(position = position_dodge(width = dodge_width), aes(size = Significant)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "orange") +
  theme_minimal() +
  labs(
    title = "OLS estimation for Political Efficacies",  # Add your plot title here
    x = "Variables",
    y = "Estimated Coefficients"
  ) +
  coord_flip() +  # Flips the axes for better visualization of terms
  scale_color_manual(values = c("darkblue", "darkred", "darkgreen"),
                     labels = c("EPE", "IPE", "OPE")) +  # Set custom colors for each model
  scale_size_manual(values = c(0.2, 0.7), guide = FALSE) +  # Adjust line width based on significance
  theme(
    legend.position = "right",  # Hide the legend for line size
    axis.text.y = element_text(size = 8)  # Adjust text size if needed
  )
```


### Interactions

```{r}

ols_ope_i <- lm(ope ~ age_num + ses + educ + sex + interest + media + social + franja + poluse + ideo_polarization + affect_polarization + social : media, data = data_w1_scores)
summary(ols_ope_i)

ols_intef_i <- lm(intef ~ age_num + ses + educ + sex + interest + media + social + franja + poluse + ideo_polarization + affect_polarization + social : media, data = data_w1_scores)
summary(ols_intef_i)

ols_extef_i <- lm(extef ~ age_num + ses + educ + sex + interest + media + social + franja + poluse + ideo_polarization + affect_polarization + social : media, data = data_w1_scores)
summary(ols_extef_i)



```

```{r echo=FALSE, message=FALSE, warning=FALSE}
plot1 <- interact_plot(model = ols_ope_i, pred = "social", modx = "media",
                       interval = TRUE, x.label = "Social Media", y.label = "OPE",
                       modx.values = "plus-minus",  # plots mean +/- 1 SD of incidental
                       modx.labels = c("Low Traditional Media", "High Traditional Media"))

plot2 <- interact_plot(model = ols_intef_i, pred = "media", modx = "social",
                       interval = TRUE, x.label = "Traditional Media", y.label = "IPE",
                       modx.values = "plus-minus",
                       modx.labels = c("Low Social Media", "High Social Media"))


# Remove the legend title

plot1 + guides(color = guide_legend(title = NULL),
               fill = guide_legend(title = NULL),
               linetype = guide_legend(title = NULL))

plot2 + guides(color = guide_legend(title = NULL),
               fill = guide_legend(title = NULL),
               linetype = guide_legend(title = NULL))




```
